Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-17 paper/paper.tex
hiding authors
LB intro
wip

\author{Maciej Pacut}
\email{maciej.pacut@univie.ac.at}
\orcid{0000-0002-6379-1490}
\affiliation{%
  \institution{Faculty of Computer Science, University of Vienna}
  \country{Austria}
}

@@

\author{Stefan Schmid} 
\email{stefan_schmid@univie.ac.at}
\affiliation{%
  \institution{Faculty of Computer Science, University of Vienna}
  \country{Austria}
}

@@

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic algorithm.
The minimum cluster capacity for the construction to remain correct is $3$.

@@

@@

  Now we estimate the cost of $\OPT(\sigma_E)$ by accounting savings from migrations.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This may allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4 (\alpha-1)}$.
  This gives us
  \begin{equation*}
    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3(\alpha - 1)\cdot \ell) \cdot \frac{\alpha}{4 (\alpha-1)}
    \geq |\sigma_E| / 4 - 4 \alpha \cdot \ell.
  \end{equation*}
  From that, we obtain $|\sigma_E| \geq 4(\OPT(\sigma_E)+\OPT(M)) + 16\alpha \cdot \ell$.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
  \end{equation*}

>>>>

\begin{comment}
\author{Maciej Pacut}
\email{maciej.pacut@univie.ac.at}
\orcid{0000-0002-6379-1490}
\affiliation{%
  \institution{Faculty of Computer Science, University of Vienna}
  \country{Austria}
}

@@

\author{Stefan Schmid} 
\email{stefan_schmid@univie.ac.at}
\affiliation{%
  \institution{Faculty of Computer Science, University of Vienna}
  \country{Austria}
}
\end{comment}

@@

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic algorithm.
It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
For $k=2$ a constant competitive algorithm exists.

@@

\pagebreak

@@

  Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This may allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4 (\alpha-1)}$.
  This gives us
  \begin{equation*}
    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3(\alpha - 1)\cdot \ell) \cdot \frac{\alpha}{4 (\alpha-1)}
    \geq |\sigma_E| / 4 - 4 \alpha \cdot \ell.
  \end{equation*}
  From that, we obtain $|\sigma_E| \geq 4(\OPT(\sigma_E)+\OPT(M)) + 16\alpha \cdot \ell$.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
  \end{equation*}
